---
title: "p8105_hw2_csc2233"
author: "Cameron Chesbrough"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## R Markdown

```{r}
subways_df = read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
subways_df = janitor::clean_names(subways_df)
subways_df_interests = select(subways_df, line, station_name, station_latitude,
                              station_longitude, route1:route11, entry,
                              vending, entrance_type, ada) 

distinct_stations = nrow(distinct(subways_df_interests, line, station_name, .keep_all = FALSE))
ada_stations = nrow(filter(subways_df_interests, ada == TRUE))
no_vending_yes_entrances = filter(subways_df_interests, entry == 'YES', vending == 'NO')
prop_no_vending_yes_entry = nrow(no_vending_yes_entrances)/nrow(subways_df_interests)

```

# Description of Subway Data

This dataset contains information about the New York subway system describing each entry/exit. This dataset contains 1868 rows and 19 columns. First I imported in the csv file. Next I used janitor to clean the data efficiently. Finally I selected the columns of importance. There are 465 distinct stations. There are 468 stations that are ADA compliant. 3.7% of stations allow entry without vending.

```{r}
mutate(subways_df_interests,
       station_line = paste(station_name,line, sep = ' '))


```


```{r}
trash_df = read_excel("202309 Trash Wheel Collection Data.xlsx",
                      sheet = 1)
trash_df = trash_df[, -c(15:16)]

trash_df = janitor::clean_names(trash_df) %>%
  mutate(sports_balls = round(sports_balls)) %>%
  mutate(sports_balls = as.integer(sports_balls))

name1 = rep("Mr_Trash_Wheel", nrow(trash_df))
Mr_trash_df = cbind(name1, trash_df)

###

trash_df2 = read_excel("202309 Trash Wheel Collection Data.xlsx",
                      sheet = 2)

trash_df2 = janitor::clean_names(trash_df2)

name2 = rep("Professor_Trash_Wheel", nrow(trash_df2))
Professor_trash_df = cbind(name2, trash_df2)

###

trash_df3 = read_excel("202309 Trash Wheel Collection Data.xlsx",
                      sheet = 4)

trash_df3 = janitor::clean_names(trash_df3)

name3 = rep("Gwynnda", nrow(trash_df3))
Gwynnda_df = cbind(name3, trash_df3)

###

all_trash_df = ()

```
